---
title: "Join microbiome"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      code: hide
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
#bibliography: mylib.bib
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
---


# Introduction


# Preparations

## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
#v.01
# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 6,
               fig.width = 5 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)
library("rmarkdown")

# Pretty outputs
library("BiocStyle")
library("ggthemes") # add ggplot2 themes
library("ggplot2")
theme_set(theme_few(base_size = 14))

# Set seed for reproducibility
set.seed(13)
# Color blind friendly pallet
cbp1 <- c("#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")

scale_fill_discrete <- function(...) {
  scale_fill_manual(..., values = cbp1)
}
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = cbp1)
}

# Set output directory
d.out <- params$d.out
rm(params)
```


## Load libraries for the session

```{r}
library(ggplot2)
library(tidyverse)
library(phyloseq)
library(vegan)
```

## Import data
```{r}
dic <- "~/IKMB/projects/skin.associations/scripts/data.dic/variables.kora.pop.tsv" %>% 
  read.delim(stringsAsFactors = F) %>% 
  mutate(Description = Displayed.description)

pop.folder <- "~/IKMB/projects/skin.associations/results/popgen/1.full.set/2.get.diversity/" 
kora.folder <- "~/IKMB/projects/skin.associations/results/kora/1.full.set/2.get.diversity/"


# Read tax

pop.tax <- "~/IKMB/projects/skin.associations/popgen/results/2.combine.metadata.microbiome/tax.rds" %>% 
  readRDS()

kora.tax <- "~/IKMB/projects/skin.associations/kora/results/2.combine.metadata.microbiome/all.tax.rds" %>% 
  readRDS()

#Read pop data

pop.pheno <- NULL
for (i in list.files(pop.folder, "i[1-9].pheno.rds")){
  pop.pheno <- bind_rows(pop.pheno,
                         paste0(pop.folder,"/",i) %>% 
                           readRDS() %>% 
                           mutate(Site = case_when(i %>% str_sub(2,2) == "1" ~ "Antecubital fossa",
                                                   i %>% str_sub(2,2) == "2" ~ "Volar forearm",
                                                    i %>% str_sub(2,2) == "3" ~ "Dorsal forearm",
                                                    i %>% str_sub(2,2) == "4" ~ "Forehead")))
}
pop.pheno <- pop.pheno %>% select(NGS_ID, Participant_ID, Site, everything())

pop.micro <- NULL
for (i in list.files(pop.folder, "i[1-9].micro.rds")){
  pop.micro <- bind_rows(pop.micro,
                         paste0(pop.folder,"/",i) %>% 
                           readRDS() %>% 
                           rownames_to_column("NGS_ID"))
}
pop.micro <- pop.micro %>% column_to_rownames("NGS_ID")
pop.micro[is.na(pop.micro)] <- 0


#Read kora data

kora.pheno <- NULL
for (i in list.files(kora.folder, "i[1-9].pheno.rds")){
  kora.pheno <- bind_rows(kora.pheno,
                         paste0(kora.folder,"/",i) %>% 
                           readRDS() %>% 
                           mutate(Site = case_when(i %>% str_sub(2,2) == "1" ~ "Antecubital fossa",
                                                   i %>% str_sub(2,2) == "2" ~ "Retroauricular fold")))
}
kora.pheno <- kora.pheno %>% select(NGS_ID, Participant_ID, Site, everything())

kora.micro <- NULL
for (i in list.files(kora.folder, "i[1-9].micro.rds")){
  kora.micro <- bind_rows(kora.micro,
                         paste0(kora.folder,"/",i) %>% 
                           readRDS() %>% 
                           rownames_to_column("NGS_ID"))
}
kora.micro <- kora.micro %>% column_to_rownames("NGS_ID")
kora.micro[is.na(kora.micro)] <- 0

```

# Make function to format data to phyloseq


```{r}
get.phyloseq <- function(micro, tax){
  

  ## Re-order columns of tax
  
  tax <- tax %>%
    right_join(data.frame(ASV = colnames(micro)), by = "ASV") 
  
  identical(tax$ASV, colnames(micro)) %>% 
    print()
  
  tax <- tax %>% 
    remove_rownames() %>% 
    select(Kingdom, Phylum, Class, Order, Family, Genus, Species, seqs) %>% 
    column_to_rownames("seqs")
    
  colnames(micro) <- rownames(tax)
  
  ## Check compatibility
  
  identical(colnames(micro), rownames(tax))
  
  # Convert to phyloseq format
  
  
  tax.ps <- tax_table(tax %>% as.matrix())
  micro.ps <- otu_table(micro, taxa_are_rows = F)
  
  # Combine data into a single object
  
  ps <- phyloseq(tax_table(tax.ps),  
                 otu_table(micro.ps, taxa_are_rows = F))
  return(ps)
}

```

# Transform in phyloseq

## Kora
```{r}
kora.ps <- get.phyloseq(kora.micro, kora.tax)
```

## Pop

```{r}
pop.ps <- get.phyloseq(pop.micro, pop.tax)
```

# Join them

```{r}
comb.ps <- merge_phyloseq(kora.ps,pop.ps)
```

# Get out of phyloseq


```{r}
comb.micro <- comb.ps %>% 
  otu_table() %>% 
  data.frame()
comb.tax <- comb.ps %>% 
  tax_table() %>% 
  data.frame(stringsAsFactors = F)
```

# Format comb.tax

```{r}
comb.tax <- comb.tax %>%
  rownames_to_column("seqs") %>% 
  mutate(ASV = 1:nrow(.) %>% 
           paste0("ASV", .)) %>% 
  select(-seqs, seqs)
```

# Format ASVs 
So ASVs combined will be ASV

```{r}
kora.tax.min <- kora.tax %>% 
  rename("ASVk" = "ASV") %>% 
  select(seqs, ASVk)

pop.tax.min <- pop.tax %>% 
  rename("ASVp" = "ASV") %>% 
  select(seqs, ASVp)

comb.tax <- comb.tax %>% 
  left_join(kora.tax.min, by = "seqs") %>% 
  left_join(pop.tax.min, by = "seqs")
```


# Format com.micro

```{r}
identical(colnames(comb.micro), comb.tax$seqs)
colnames(comb.micro) <- comb.tax$ASV 
identical(colnames(comb.micro), comb.tax$ASV)
```

# Sync with metadata

## Select variables that are common

```{r}
dic <- dic %>% 
  rename("uDescription" = "Description",
         "varname" = "Variable") %>% 
  bind_rows(data.frame(Cohort = c(rep("KORA FF4",2),rep("PopGen",2)),
                       uDescription = rep(c("NGS_ID", "Participant_ID"),2),
                       varname = rep(c("NGS_ID", "Participant_ID"),2)))
common.variables <- dic %>%
  group_by(uDescription) %>% 
  mutate(n = n()) %>% 
  filter(n == 2) %>% 
  arrange(uDescription) %>% 
  ungroup
```

## Standardize kora meta and dic
```{r}
# Select variables to consider
in.kora.dic <- dic %>% 
  filter(Cohort == "KORA FF4" ) %>% 
  inner_join(common.variables %>% filter(Cohort == "KORA FF4") %>% select(varname), by = "varname")
in.kora.pheno <- kora.pheno %>% 
  select(match((in.kora.dic %>% pull(varname)), colnames(.)))
identical(colnames(in.kora.pheno), in.kora.dic$varname)
#Edit dic and meta
in.kora.dic <- in.kora.dic %>% 
  mutate(varname = uDescription %>% make.names(unique = T)) %>% 
  mutate(Description = uDescription) %>% 
  select(varname, Group,Description)
colnames(in.kora.pheno) <- in.kora.dic$varname
identical(colnames(in.kora.pheno), in.kora.dic$varname)
```

## Standardize pop meta and dic
```{r}
# Select variables to consider
in.pop.dic <- dic %>% 
  filter(Cohort == "PopGen" ) %>% 
  inner_join(common.variables %>% filter(Cohort == "PopGen") %>% select(varname), by = "varname")
in.pop.pheno <- pop.pheno %>% 
  select(match((in.pop.dic %>% pull(varname)), colnames(.)))
identical(colnames(in.pop.pheno), in.pop.dic$varname)
#Edit dic and meta
in.pop.dic <- in.pop.dic %>% 
  mutate(varname = uDescription %>% make.names(unique = T)) %>% 
  mutate(Description = uDescription) %>% 
  select(varname, Group,Description)
colnames(in.pop.pheno) <- in.pop.dic$varname
identical(colnames(in.pop.pheno), in.pop.dic$varname)
```

## Join phenotypic data

```{r}
comb.pheno <- bind_rows(in.kora.pheno %>% 
                          mutate(Cohort = "KORA FF4"),
                        in.pop.pheno %>% 
                          mutate(Cohort = "PopGen")) %>% 
  group_by(Cohort)
```


# Sync meta with micro

```{r}
comb.pheno <- comb.micro %>% 
  rownames_to_column("NGS_ID") %>% 
  select(NGS_ID) %>% 
  left_join(comb.pheno, by = "NGS_ID")
```


## Overview of the data

```{r}
for (i in setdiff(colnames(comb.pheno), c("NGS_ID", "Participant_ID"))){
  
  if(is.numeric(comb.pheno[,i])){
    g <- ggplot(comb.pheno, aes_string(x = i)) +
      geom_histogram() + 
      facet_wrap(~Cohort) +
      labs(title = i)}else{
        g <- ggplot(comb.pheno, aes_string(x = i)) +
          geom_bar() + 
          facet_wrap(~Cohort)  +
          labs(title = i)}
  print(g)
}
```


## Check if data fits

```{r}
identical(rownames(comb.micro), comb.pheno$NGS_ID) 
identical(colnames(comb.micro), comb.tax$ASV) 
```


# Make micro rarefied

# Micro rarefied
```{r}
min.count <- 5000
micro.r <- rrarefy(comb.micro, min.count) %>% 
  data.frame()
micro.r <-  micro.r[, colSums(micro.r) > 0]
```

# Create master tax

This will include all ASVs from the pop and kora

## Creat unique taxonomy

```{r}
tax.unique <- bind_rows(pop.tax %>% select(-ASV),
                        kora.tax %>% select(-ASV)) %>% 
  filter(!duplicated(seqs))
```
## Join the ASV names

```{r}
anti.kora.tax <- anti_join(kora.tax %>% 
                             select(ASV, seqs) %>% 
                             rename("ASVk" = "ASV"), 
                           comb.tax %>% select(seqs),
                           by = "seqs")
anti.pop.tax <- anti_join(pop.tax %>% 
                             select(ASV, seqs) %>% 
                            rename("ASVp" = "ASV"),
                          comb.tax %>%  select(seqs),
                           by = "seqs")
anti.tax <- full_join(anti.kora.tax, anti.pop.tax, by = "seqs")

#Add ASV count

anti.tax <- anti.tax %>% 
  mutate(ASV = 1:nrow(anti.tax) + nrow(comb.tax),
         ASV = paste0("ASV", ASV)) %>% 
  #add taxonomy
  left_join(tax.unique, by = "seqs")

```

## Master tax

```{r}
master.tax <-bind_rows(comb.tax,
                       anti.tax)
(duplicated(master.tax$seqs) %>% sum) == 0
```


# Write outputs

```{r}
"meta.rds" %>%
  paste0(d.out, "/", .) %>% 
  saveRDS(comb.pheno, . )

"micro.rds" %>%
  paste0(d.out, "/", .) %>% 
  saveRDS(comb.micro, .)

"micro.r.rds" %>%
  paste0(d.out, "/", .) %>% 
  saveRDS(micro.r, .)

"tax.rds" %>%
  paste0(d.out, "/", .) %>% 
  saveRDS(comb.tax, .)

"master.tax.rds" %>%
  paste0(d.out, "/", .) %>% 
  saveRDS(master.tax, .)
```


# Conclusion

# Session information

```{r}
sessionInfo()
```
