---
title: "Make summary table"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      code: hide
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
#bibliography: mylib.bib
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
---


# Introduction


# Preparations

## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
#v.01
# Set knit global options
library("knitr")
options(digits = 1, width = 120, qwraps2_markup = "markdown")
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 6,
               fig.width = 5 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)
library("rmarkdown")

# Pretty outputs
library("BiocStyle")
library("ggthemes") # add ggplot2 themes
library("ggplot2")
theme_set(theme_few(base_size = 14))

# Set seed for reproducibility
set.seed(13)
# Color blind friendly pallet
cbp1 <- c("#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")

scale_fill_discrete <- function(...) {
  scale_fill_manual(..., values = cbp1)
}
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = cbp1)
}

# Set output directory
d.out <- params$d.out
rm(params)
```


## Load libraries for the session

```{r}
library(ggplot2)
library(tidyverse)
library(qwraps2)
```

## Import data
```{r}
dic <- "/home/lsilva/IKMB/projects/skin.associations/results/analysis/2.summary.tables/1.host.trait.set/variables.kora.pop.tsv" %>% 
  read.delim(stringsAsFactors = F)

pop.data.folder <- "/home/lsilva/IKMB/projects/skin.associations/results/popgen/2.host.trait.set"
kora.data.folder <- "/home/lsilva/IKMB/projects/skin.associations/results/kora/2.host.trait.set"

# get data folder
pop.folder <- paste0(pop.data.folder, "/", "1.filter.samples.and.metadata")
kora.folder <- paste0(kora.data.folder, "/", "1.filter.samples.and.metadata")

#Read pop data
fin <- data.frame(index = c("i1","i2","i3", "i4"), Site = c("Antecubital fossa","Volar forearm", "Dorsal forearm",  "Forehead"))


pop.pheno <- NULL
for (i in 1:nrow(fin)){
  pop.pheno <- paste0(pop.data.folder, "/2.get.diversity/",
                      fin$index[i],
                      ".pheno.rds") %>% 
    readRDS() %>% 
    mutate(Site = fin$Site[i]) %>% 
    bind_rows(pop.pheno,.)
}


# Read kora data
#Read pop data
fin <- data.frame(index = c("i1","i2"), Site = c("Antecubital fossa","Retroauricular fold"))


kora.pheno <- NULL
for (i in 1:nrow(fin)){
  kora.pheno <- paste0(kora.data.folder, "/2.get.diversity/",
                       fin$index[i],
                       ".pheno.rds") %>% 
    readRDS() %>% 
    mutate(Site = fin$Site[i]) %>% 
    bind_rows(kora.pheno,.)
}
kora.dic.values <- "pheno.dic.values.all.gold.rds" %>% 
  paste0(kora.folder, "/", .) %>% 
  readRDS(.)
```

# Get common table

# Sync with metadata

## Select variables that are common

```{r}
dic <- dic %>% 
  # NEW feature!!!
  bind_rows(data.frame(Variable = rep("Participant_ID",2), 
                       Displayed.description = rep("Participant_ID",2),
                       Variable = rep("Participant_ID",2),
                       Cohort = c("KORA FF4", "PopGen"))) %>% 
  rename("uDescription" = "Displayed.description",
         "varname" = "Variable")

common.variables <- dic %>%
  group_by(uDescription) %>% 
  mutate(n = n()) %>% 
  arrange(uDescription) %>% 
  ungroup
```

## Standardize kora meta and dic
This is necessary because names in each phenotype datasets are cohort-specific.
```{r}
# Select variables to consider
in.kora.dic <- dic %>% 
  filter(Cohort == "KORA FF4" ) %>% 
  inner_join(common.variables %>%
               filter(Cohort == "KORA FF4") %>%
               select(varname), by = "varname")

in.kora.pheno <- kora.pheno %>% 
  select(match((in.kora.dic %>% pull(varname)), colnames(.)))

identical(colnames(in.kora.pheno), in.kora.dic$varname)

#Edit dic and meta
in.kora.dic <- in.kora.dic %>% 
  mutate(varname = uDescription %>% make.names(unique = T)) %>% 
  mutate(Description = uDescription) %>% 
  select(varname, Group,Description)
colnames(in.kora.pheno) <- in.kora.dic$varname

identical(colnames(in.kora.pheno), in.kora.dic$varname)
```

## Standardize pop meta and dic
This is necessary because names in each phenotype datasets are cohort-specific.
```{r}
# Select variables to consider
in.pop.dic <- dic %>% 
  filter(Cohort == "PopGen" ) %>% 
  inner_join(common.variables %>% 
               filter(Cohort == "PopGen") %>% 
               select(varname), by = "varname")

in.pop.pheno <- pop.pheno %>% 
  select(match((in.pop.dic %>% pull(varname)), colnames(.)))

identical(colnames(in.pop.pheno), in.pop.dic$varname)

#Edit dic and meta
in.pop.dic <- in.pop.dic %>% 
  mutate(varname = uDescription %>% make.names(unique = T)) %>% 
  mutate(Description = uDescription) %>% 
  select(varname, Group,Description)
colnames(in.pop.pheno) <- in.pop.dic$varname

identical(colnames(in.pop.pheno), in.pop.dic$varname)
```

## Join phenotypic data

```{r}
comb.pheno <- bind_rows(in.kora.pheno %>% 
                          mutate(Cohort = "KORA FF4"), 
                        #Keep only samples that are antecubital fossa
                        #filter(Site  == "Antecubital fossa"),
                        in.pop.pheno %>% 
                          mutate(Cohort = "PopGen")) %>% 
  group_by(Cohort) %>% 
  # Add new name of the Site
  mutate(Site = if_else(Site == "Behind ear",
                        "Retroauricular fold",
                        Site)) %>% 
  #update names of wash and lotion
  rename("Wash" = "Wash...24h.",
         "Lotion" = "Lotion.cream...24h.") %>% 
  mutate(Sport.in.summer..h.week. = if_else(is.na(Sport.in.summer..h.week.),
                                            0,
                                            Sport.in.summer..h.week.),
         Regular.sports = if_else(is.na(Regular.sports),
                                  0L,
                                  Regular.sports)) %>% 
  mutate(Sex = str_to_sentence(Sex))

```

# Make table

## Participant data
```{r, results="asis"}
comb.pheno.participant <- comb.pheno %>% 
  select(-Site, -Wash, -Lotion) %>% 
  unique() %>% 
  mutate()

our_summary1 <- list(
  "Age/BMI/Sex" = list("Age (years)" = ~ qwraps2::median_iqr(.data$Age..y., digits = 0),
                       "Sex:female" = ~ n_perc0(.data$Sex == "Female"),
                       "BMI" = ~ qwraps2::median_iqr(.data$BMI..kg.m2., digits = 0)),
  "Diet" = list("Carbohydrate (g/d)" = ~ qwraps2::median_iqr(.data$Carbohydrate..g.d., digits = 0),
                "Energy (kcal/d)" = ~ qwraps2::median_iqr(.data$Energy..kcal.d., digits = 0),
                "Fat (g/d)" = ~ qwraps2::median_iqr(.data$Fat..g.d., digits = 0),
                "Fiber (g/d)" = ~ qwraps2::median_iqr(.data$Fiber..g.d., digits = 0),
                "Protein (g/d)" = ~ qwraps2::median_iqr(.data$Protein..g.d., digits = 0)),
  "Systemic antibiotics" = list("Antibiotics (< 6-8w):yes" = ~ n_perc0(.data$Antibiotics...6.8w. == "yes")),
  "Smoking" = list("Smoking:ex-smoker" = ~ n_perc0(.data$Smoking %in% c("Ex-smoker", "ex smoker")),
                   "Smoking:non-smoker" = ~ n_perc0(.data$Smoking %in% c("never smoked", "Non-smoker")),
                   "Smoking:smoker" = ~ n_perc0(.data$Smoking %in% c("smoker", "Smoker"))),
  "Alcohol consumption" = list("Alcohol (g/d)" = ~ qwraps2::median_iqr(.data$Alcohol..g.d., digits = 0)),
  "Physical activity" = list("Regular sports*" = ~ n_perc0(.data$Regular.sports == "1"),
                             "Sport in summer (h)*" = ~ qwraps2::median_iqr(.data$Sport.in.summer..h.week.)))


whole <- summary_table(comb.pheno.participant, our_summary1)
#print(whole, markup = "latex")
whole
```

# Session information

```{r}
sessionInfo()
```
