---
title: "Visualize diversity patterns of PopGen skin microbiome"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
#bibliography: mylib.bib
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
---


# Introduction

Here I will visually explore the diversity of PopGen skin microbiome. I will work with the pre-computed (see folder before) diversity measures. I want to visualize obvious patterns between the diversity measures and obvious phenotypes.

# Preparations

## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 5,
               fig.width = 4 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)
library("rmarkdown")

# Pretty outputs
library("BiocStyle")
library("ggthemes") # add ggplot2 themes
library("ggplot2")
theme_set(theme_few(base_size = 14))

# Set seed for reproducibility
set.seed(13)

# Set output directory
d.out <- params$d.out
rm(params)
```


## Load libraries for the session

```{r}
library(ggplot2)
library(tidyverse)
library(vegan)
library(reshape2)
```

## Import data
```{r}
# Input folders
d.in.diversity <- "/home/lsilva/IKMB/projects/skin.associations/results/analysis/1.combine.microbiome/1.join.microbiome/"

microenv <- "/home/lsilva/IKMB/projects/skin.associations/scripts/data.dic/skin.microenv.tsv" %>% 
  read.delim(stringsAsFactors = F)

meta <- "meta.rds" %>% 
  paste0(d.in.diversity, "/", .) %>% 
  readRDS() %>% 
  left_join(microenv, by = "Site") %>% 
  mutate(Site = if_else(Site == "Antecubital fossa" &
                          Cohort == "KORA FF4",
                        "Antecubital fossa (K)",
                        Site)) %>% 
  mutate(Site = if_else(Site == "Antecubital fossa" &
                          Cohort == "PopGen",
                        "Antecubital fossa (P)",
                        Site))

micro.r <- "micro.r.rds" %>% 
  paste0(d.in.diversity, "/", .) %>% 
  readRDS()
```


# Get distances

```{r}
micro.dis <- micro.r %>%
  vegdist("bray")
```

# Beta-diversity

Let's have a look at the beta diversity

## Calculate mds
```{r}
mds <- micro.dis %>% 
  metaMDS(.,autotransform = F, trymax = 50)
```

## Visualize distance of samples
```{r}

to.plot.beta <- mds$points %>%
  data.frame() %>% 
  rownames_to_column("NGS_ID") %>% 
  left_join(meta) %>%
  group_by(Participant_ID)


ggpubr::ggscatterhist(
  to.plot.beta %>% 
    mutate(Site = factor(Site, levels = c("Dorsal forearm",
                                          "Volar forearm",
                                          "Antecubital fossa (P)",
                                          "Antecubital fossa (K)",
                                          "Forehead",
                                          "Retroauricular fold"))),
  x = "MDS1", y = "MDS2",
  color = "Site",
  alpha = 0.4,
  margin.params = list(fill = "Site", color = "darkgrey", size = 0.2),
  margin.plot = "boxplot",
  ggtheme = theme_few(),
)

```

# Save

```{r}
"to.plot.beta.rds" %>%  
  paste0(d.out, "/", .) %>% 
  saveRDS(to.plot.beta, .)

```


# Session information

```{r}
sessionInfo()
```
