---
title: "Visualize diversity patterns of PopGen skin microbiome"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
#bibliography: mylib.bib
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
---


# Introduction

Here I will visually explore the diversity of PopGen skin microbiome. I will work with the pre-computed (see folder before) diversity measures. I want to visualize obvious patterns between the diversity measures and obvious phenotypes.

# Preparations

## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 5,
               fig.width = 4 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)
library("rmarkdown")

# Pretty outputs
library("BiocStyle")
library("ggthemes") # add ggplot2 themes
library("ggplot2")
theme_set(theme_few(base_size = 14))

# Set seed for reproducibility
set.seed(13)

# Set output directory
d.out <- params$d.out
rm(params)
```


## Load libraries for the session

```{r}
library(ggplot2)
library(tidyverse)
library(vegan)
library(reshape2)
library(ape)
library(RColorBrewer)
```

## Import data
```{r}
# Input folders
d.in.diversity <- "/home/lsilva/IKMB/projects/skin.associations/results/analysis/1.combine.microbiome/1.join.microbiome/"

microenv <- "/home/lsilva/IKMB/projects/skin.associations/scripts/data.dic/skin.microenv.tsv" %>% 
  read.delim(stringsAsFactors = F)

meta <- "meta.rds" %>% 
  paste0(d.in.diversity, "/", .) %>% 
  readRDS() %>% 
  left_join(microenv, by = "Site") %>% 
  mutate(Site = if_else(Site == "Antecubital fossa" &
                          Cohort == "KORA FF4",
                        "Antecubital fossa (K)",
                        Site)) %>% 
  mutate(Site = if_else(Site == "Antecubital fossa" &
                          Cohort == "PopGen",
                        "Antecubital fossa (P)",
                        Site))

micro.r <- "micro.r.rds" %>% 
  paste0(d.in.diversity, "/", .) %>% 
  readRDS()

micro <- "micro.rds" %>% 
  paste0(d.in.diversity, "/", .) %>% 
  readRDS()

# Retrieve sequencing runs

pop.run <- "/home/lsilva/IKMB/projects/skin.associations/popgen/metadata/runs.files/runs.files.tsv" %>% 
  read.delim(header = T) # Includes popgen and and some kora ff4

kora.run <- readRDS("/home/lsilva/IKMB/projects/skin.associations/kora/results/2.combine.metadata.microbiome/all.meta.rds") %>% 
  select(NGS_ID, run) %>% 
  rename("kora.run" = "run")
```


# Function

Add quickRareCurve from Dave Clark @ https://dave-clark.github.io/post/speeding-up-rarefaction-curves-for-microbial-community-ecology/
```{r}
quickRareCurve <- function (x, step = 1, sample, xlab = "Sample Size",
                            ylab = "Species", label = TRUE, col, lty, max.cores = T, nCores = 1, ...)
{
  require(parallel)
  x <- as.matrix(x)
  if (!identical(all.equal(x, round(x)), TRUE))
    stop("function accepts only integers (counts)")
  if (missing(col))
    col <- par("col")
  if (missing(lty))
    lty <- par("lty")
  tot <- rowSums(x) # calculates library sizes
  S <- specnumber(x) # calculates n species for each sample
  if (any(S <= 0)) {
    message("empty rows removed")
    x <- x[S > 0, , drop = FALSE]
    tot <- tot[S > 0]
    S <- S[S > 0]
  } # removes any empty rows
  nr <- nrow(x) # number of samples
  col <- rep(col, length.out = nr)
  lty <- rep(lty, length.out = nr)
  # parallel mclapply
  # set number of cores
  mc <- getOption("mc.cores", ifelse(max.cores, detectCores(), nCores))
  message(paste("Using ", mc, " cores"))
  out <- mclapply(seq_len(nr), mc.cores = mc, function(i) {
    n <- seq(1, tot[i], by = step)
    if (n[length(n)] != tot[i])
      n <- c(n, tot[i])
    drop(rarefy(x[i, ], n))
  })
  Nmax <- sapply(out, function(x) max(attr(x, "Subsample")))
  Smax <- sapply(out, max)
  plot(c(1, max(Nmax)), c(1, max(Smax)), xlab = xlab, ylab = ylab,
       type = "n", ...)
  if (!missing(sample)) {
    abline(v = sample)
    rare <- sapply(out, function(z) approx(x = attr(z, "Subsample"),
                                           y = z, xout = sample, rule = 1)$y)
    abline(h = rare, lwd = 0.5)
  }
  for (ln in seq_along(out)) {
    N <- attr(out[[ln]], "Subsample")
    lines(N, out[[ln]], col = col[ln], lty = lty[ln], ...)
  }
  if (label) {
    ordilabel(cbind(tot, S), labels = rownames(x), ...)
  }
  invisible(out)
}
```

# Get distances

```{r}
micro.dis <- micro.r %>%
  vegdist("bray")
```

# Beta-diversity

Let's have a look at the beta diversity

## Calculate pcoa
```{r}
pcoa <- micro.dis %>% 
  pcoa()
pcoa$values[1:5,]
biplot.pcoa(pcoa)
```

## Visualize distance of samples
```{r}

to.plot.beta <- pcoa$vectors[,1:2] %>%
  data.frame() %>% 
  rownames_to_column("NGS_ID") %>% 
  left_join(meta) %>%
  group_by(Participant_ID)


ggpubr::ggscatterhist(
  to.plot.beta %>% 
    mutate(Site = factor(Site, levels = c("Dorsal forearm",
                                          "Volar forearm",
                                          "Antecubital fossa (P)",
                                          "Antecubital fossa (K)",
                                          "Forehead",
                                          "Retroauricular fold"))),
  x = "Axis.1", y = "Axis.2",
  color = "Site",
  alpha = 0.4,
  margin.params = list(fill = "Site", color = "darkgrey", size = 0.2),
  margin.plot = "boxplot",
  ggtheme = theme_few(),
)

```

# Save

```{r}
"to.plot.beta.rds" %>%  
  paste0(d.out, "/", .) %>% 
  saveRDS(to.plot.beta, .)

"pcoa.data.rds" %>%  
  paste0(d.out, "/", .) %>% 
  saveRDS(pcoa$values, .)

```

# Plot sequencing runs

## Edit information
```{r}
# set colnames
pop.run <- pop.run %>% 
  rownames_to_column("V1") %>% 
  select(V1) %>% 
  separate(V1,into = c("pop.run", "garbage","NGS_ID"), sep = "/|:") %>% 
  filter(NGS_ID != "input") %>% 
  select(-garbage) %>% 
  select(pop.run, NGS_ID) %>% 
  #Remove repeated run
  filter(pop.run != "H19847") # It repeats with H19847A. It is basically similar info, but Malte repeated H19847, making H19847A because he could not access some info in the original folder. 
```

## Plot

```{r}
to.plot.beta <- to.plot.beta %>% 
  #add pop.run
  left_join(pop.run, by = "NGS_ID") %>% 
  # add kora.run
  left_join(kora.run, by = "NGS_ID") %>% 
  # unify runs
  mutate(Run = case_when(Cohort == "PopGen" ~ pop.run,
                         Cohort == "KORA FF4" ~ kora.run)) %>% 
  #code runs
  ungroup() %>% 
  mutate(run.id = Run %>%
           factor() %>% 
           as.numeric %>% 
           as.character() %>% 
           str_pad(2, "left", pad = "0"))

ncol <- to.plot.beta$Run %>% unique() %>% length()
mycolors <- colorRampPalette(brewer.pal(8, "Paired"))(ncol)

batch.plot <- ggplot(to.plot.beta, aes(x = Axis.1, y = Axis.2, group = run.id)) +
  geom_point(aes(col = run.id), alpha = 0.7) +
  facet_wrap(~ Cohort + Site, scales = "free") +
  stat_ellipse(aes(col = run.id)) +
  labs(color = "Sequencing run",
       x = paste0("Axis 1 (", 
                  pcoa$values$Relative_eig[1] %>% 
                    `*`(100) %>% 
                    round(2),
                  "%)"),
       y = paste0("Axis 2 (", 
                  pcoa$values$Relative_eig[2] %>% 
                    `*`(100) %>% 
                    round(2),
                  "%)")) +
  theme(legend.position = "bottom") +
  scale_color_manual(values = mycolors) +
  guides(color=guide_legend(ncol=11))
```


## Summarize

```{r}
to.plot.beta %>% select(Cohort, run.id) %>% 
  unique() %>% 
  group_by(Cohort) %>% 
  summarize(n = n())


to.plot.beta %>% select(Cohort, run.id, Site) %>% 
  unique() %>% 
  View()
```


# Get number of sequences per run

```{r}
sequences.plot <- micro %>% rowSums() %>% 
  data.frame(NGS_ID = names(.), total.seqs = .) %>% 
  left_join(to.plot.beta %>% 
              select(NGS_ID, run.id, Site, Cohort) %>% 
              unique(), by = "NGS_ID") %>% 
  ggplot(aes(y = total.seqs, x = run.id)) +
  geom_boxplot(aes(color = run.id)) +
  geom_jitter(aes(color = run.id), alpha = 0.3, size = 0.4)+
  scale_color_manual(values = mycolors) +
  labs(x = "Sequencing Run", y = "Total sequences") +
  theme(legend.position = "none") +
  ggpubr::rotate_x_text() +
  scale_y_log10()
sequences.plot
```

# Get rarefaction curves


```{r}

curves <- quickRareCurve(micro[,], step = 100, max.cores = F, nCores = 6 ) 
names(curves) <- rownames(micro[,])


to.plot <- NULL
for (i in names(curves)){
  subsample <- curves[[i]]
  df <- data.frame(ASV = subsample %>% as.numeric, 
                   index = subsample %>% 
                     names() %>% 
                     gsub("N", "",.) %>% as.numeric(), NGS_ID = i)
  to.plot <- bind_rows(to.plot, df)
}

to.plot <- to.plot %>% 
  left_join(to.plot.beta %>% 
              select(NGS_ID, run.id, Site, Cohort) %>% 
              unique(), by = "NGS_ID")


rarecurve <- ggplot(to.plot %>%  filter(index < 10000) , aes(index, ASV, group = NGS_ID)) +
  geom_line(alpha = 0.1) +
  geom_vline(xintercept = 5000, linetype = 2) +
  facet_wrap(~ Site) +
  theme(legend.position = "none") +
  labs(x = "Sample size")
```

# Make final plot

```{r, fig.height=10, fig.width=10}

gbc <- ggpubr::ggarrange(sequences.plot, rarecurve, labels = c("A", "B"), widths = c(1,2))
ggpubr::ggarrange(gbc, batch.plot , labels = c("", "C"), nrow =2, heights = c(1,1.7))
```



# Session information

```{r}
sessionInfo()
```
