---
title: "Filter microbiome data"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
#bibliography: mylib.bib
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
---



# Preparations

## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 5,
               fig.width = 4 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)
library("rmarkdown")

# Pretty outputs
library("BiocStyle")
library("ggthemes") # add ggplot2 themes
library("ggplot2")
theme_set(theme_few(base_size = 18))

# Set seed for reproducibility
set.seed(222)
d.out <- params$d.out
rm(params)
```


## Load libraries for the session

```{r}
library(ggplot2)
library(tidyverse)
library(DataExplorer)
```

## Import data
```{r}
d.in <- "/home/lsilva/IKMB/projects/skin.associations/kora/results/2.combine.metadata.microbiome/"
meta <- "samples.meta.rds" %>%
  paste0(d.in, "/",. ) %>%
  readRDS()
micro <- "samples.micro.rds" %>%
  paste0(d.in, "/",. ) %>%
  readRDS()
tax <- "samples.tax.rds" %>%
  paste0(d.in, "/",. ) %>%
  readRDS()
meta.dic <- "all.dic.rds" %>%
  paste0(d.in, "/",. ) %>%
  readRDS()
meta.dic.values <- "all.dic.values.rds" %>%
  paste0(d.in, "/",. ) %>%
  readRDS()
```

# Introduction

I will filter the microbiome data to keep only samples with minimum sequencing depth of 10,000 sequences.

# Set minimal sequencing depth for microbiome data

## Microbiome table

### Remove samples with low depth
```{r}
dim(micro)
micro <- micro[rowSums(micro) > 10000,]
dim(micro)
```

### Remove possibly empty ASVs

```{r}
micro <- micro[,colSums(micro) > 0]
```

## Update the other tables

### Meta
```{r}
meta <- rownames(micro) %>%
  data.frame(NGS_ID = ., stringsAsFactors = F) %>%
  left_join(meta)
```

### Tax

```{r}
tax <- colnames(micro) %>%
  data.frame(ASV = ., stringsAsFactors = F) %>%
  left_join(tax)
```


```{r}
"samples.meta.f.rds" %>%
  paste0(d.out, "/", .) %>%
  saveRDS(meta, .)
"samples.micro.f.rds" %>%
  paste0(d.out, "/", .) %>%
  saveRDS(micro, .)
"samples.tax.f.rds" %>%
  paste0(d.out, "/", .) %>%
  saveRDS(tax, .)
"samples.meta.dic.f.rds" %>%
  paste0(d.out, "/", .) %>%
  saveRDS(meta.dic, .)
"samples.meta.dic.values.f.rds" %>%
  paste0(d.out, "/", .) %>%
  saveRDS(meta.dic.values, .)
```


# Session information

```{r}
sessionInfo()
```
