---
title: "Filter metadata"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
#bibliography: mylib.bib
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
---



# Preparations

## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 5,
               fig.width = 4 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)
library("rmarkdown")

# Pretty outputs
library("BiocStyle")
library("ggthemes") # add ggplot2 themes
library("ggplot2")
library(ICC)
library(vcd)

theme_set(theme_few(base_size = 18))

# Set seed for reproducibility
set.seed(222)
d.out <- params$d.out
rm(params)
```


## Load libraries for the session

```{r}
library(ggplot2)
library(tidyverse)
library(finalfit)
library(janitor)
library(DataExplorer)
```

## Import data
```{r}
d.in <- d.out

meta <- "samples.meta.f.rds" %>%
  paste0(d.out, "/", .) %>%
  readRDS()
micro <- "samples.micro.f.rds" %>%
  paste0(d.out, "/", .) %>%
  readRDS()
tax <- "samples.tax.f.rds" %>%
  paste0(d.out, "/", .) %>%
  readRDS()
meta.dic <- "samples.meta.dic.f.rds" %>%
  paste0(d.out, "/", .) %>%
  readRDS()
meta.dic.values <-"samples.meta.dic.values.f.rds" %>%
  paste0(d.out, "/", .) %>%
  readRDS()
```


# Introduction

I will filter the microbiome and the metadata to to have a gold data.

# Keep only phenotypic data

```{r}
non_pheno <- meta.dic %>%
  # Get non phenotypic variables
  filter(Group %in% c("Study", "16S Sequencing", "Sample")) %>%
  # Exclude from the list important variables for ID and eventual phenotypic
  filter(!Variable %in% c("NGS_ID", "Site", "lhr501x_v3", "Probennr")) %>%
  pull(Variable)

pheno.dic <- meta.dic %>%
  filter(!Variable %in% non_pheno)
pheno <- meta %>%
  select(which(colnames(.) %in% pheno.dic$varname))

# confirming
identical(colnames(pheno), pheno.dic$varname)
#cleaning
rm(non_pheno)
```


# Exlude participants with local antibiotic or corticosteroid treatment within 7 days

```{r}
dim(pheno)
pheno <- pheno %>%
  filter(Hyhaottaoystwaosecftp7dImapAbtle.u3pd_04_1 == 0) %>% 
  select(-Hyhaottaoystwaosecftp7dImapAbtle.u3pd_04_1)
pheno.dic <- pheno.dic %>% 
  filter(varname != "Hyhaottaoystwaosecftp7dImapAbtle.u3pd_04_1")
dim(pheno)
```

# Exclude participants with anormalities when collecting samples

```{r}
dim(pheno)
pheno <- pheno %>%
  filter(is.na(u3pd_08)) %>% 
  select(-u3pd_08)
pheno.dic <- pheno.dic %>% 
  filter(varname != "u3pd_08")
dim(pheno)
```


# Get Gold standards

```{r}
pheno.gold <- pheno
pheno.dic <- pheno.dic %>%
  right_join(pheno %>%
               colnames() %>% 
               data.frame(varname = .),
             by = "varname")
```

### Get dictionary
```{r}
pheno.gold.dic <- colnames(pheno.gold) %>%
  data.frame(varname = .) %>%
  left_join(pheno.dic, by = "varname")
```

```{r}
pheno.gold.dic.values <- meta.dic.values %>%
  filter(varname %in% pheno.dic$varname)
```

## Keep microbiome data of selected samples

## Select participants
```{r}
micro.gold <- micro[match(pheno.gold$NGS_ID, rownames(micro)),]
identical(pheno.gold$NGS_ID, rownames(micro.gold))
```


## Select ASVs that are not empty
```{r}
micro.gold <- micro.gold[,colSums(micro.gold) > 0]
```

## Match tax

```{r}
tax.gold <- colnames(micro.gold) %>%
  data.frame(ASV = .) %>%
  left_join(tax, by = "ASV")
identical(colnames(micro.gold), tax.gold$ASV)
```

# Save outputs

Gold

```{r}
"pheno.all.gold.rds" %>%
  paste0(d.out, "/", .) %>%
  saveRDS(pheno.gold, .)
"micro.all.gold.rds" %>%
  paste0(d.out, "/", .) %>%
  saveRDS(micro.gold, .)
"tax.all.gold.rds" %>%
  paste0(d.out, "/", .) %>%
  saveRDS(tax.gold, .)
"pheno.dic.all.gold.rds" %>%
  paste0(d.out, "/", .) %>%
  saveRDS(pheno.gold.dic, .)
"pheno.dic.values.all.gold.rds" %>%
  paste0(d.out, "/", .) %>%
  saveRDS(pheno.gold.dic.values, .)
```

# Session information

```{r}
sessionInfo()
```

