---
title: "Filter metadata"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      code: hide
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
#bibliography: mylib.bib
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
---


# Introduction
I will filter the microbiome and the metadata to have a small, **GOLD**, selected set of samples where phenotype data is complete.


# Preparations

## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
#v.01
# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 6,
               fig.width = 5 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)
library("rmarkdown")

# Pretty outputs
library("BiocStyle")
library("ggthemes") # add ggplot2 themes
library("ggplot2")
theme_set(theme_few(base_size = 14))

# Set seed for reproducibility
set.seed(13)
# Color blind friendly pallet
cbp1 <- c("#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")

scale_fill_discrete <- function(...) {
   scale_fill_manual(..., values = cbp1)
}
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = cbp1)
}

# Set output directory
d.out <- params$d.out
rm(params)
```


## Load libraries for the session

```{r}
library(ggplot2)
library(tidyverse)

```

## Load libraries for the session

```{r}
library(ggplot2)
library(tidyverse)
library(finalfit)
library(janitor)
library(DataExplorer)
```


## Import data
```{r}
d.in <- "/home/lsilva/IKMB/projects/skin.associations/results/kora/1.full.set/1.filter.samples.and.metadata"
meta <- "pheno.all.gold.rds" %>%
  paste0(d.in, "/", .) %>%
  readRDS()
micro <- "micro.all.gold.rds" %>%
  paste0(d.in, "/", .) %>%
  readRDS()
tax <- "tax.all.gold.rds" %>%
  paste0(d.in, "/", .) %>%
  readRDS()
meta.dic <- "pheno.dic.all.gold.rds" %>%
  paste0(d.in, "/", .) %>%
  readRDS()
meta.dic.values <-"samples.meta.dic.values.f.rds" %>%
  paste0(d.in, "/", .) %>%
  readRDS()
```


# Introduction

I will filter the microbiome and the metadata to have a small, **GOLD**, selected set of samples where phenotype data is complete.

# Rename
```{r}
pheno <- meta
pheno.dic <- meta.dic
```


# Exlude participants with underreported diet info

This step removes participants that do not have diet information.
```{r}
dim(pheno)
pheno <- pheno %>%
  filter(Under.reporter.u3v_under_rep == 0) %>% 
  select(-Under.reporter.u3v_under_rep)
pheno.dic <- pheno.dic %>% 
  filter(varname != "Under.reporter.u3v_under_rep")
dim(pheno)
```


# Select variables that match with PopGen


```{r}
dim(pheno)
pheno <- pheno %>%
  select(NGS_ID,
         Participant_ID,
         Site,
         Agetothedeadlin.u3talter,
         gender.u3csex,
         BMI.u3tbmi,
         Crbhydrtsabsrbb.u3v_zk,
         Energy.u3v_gcal,
         Fat.u3v_zf,
         Fiber.u3v_zb,
         Protein.protein.u3v_ze,
         Ddytkantbtitl2m.u3sf_01,
         smoking.u3tcigsmk,
         Alcohol.ethanol.u3v_za,
         cream,
         wash)
pheno.dic <- pheno.dic %>% 
  filter(varname %in% c("NGS_ID",
                        "Participant_ID",
                        "Site",
                        "Agetothedeadlin.u3talter",
                        "gender.u3csex",
                        "BMI.u3tbmi",
                        "Crbhydrtsabsrbb.u3v_zk",
                        "Energy.u3v_gcal",
                        "Fat.u3v_zf",
                        "Fiber.u3v_zb",
                        "Protein.protein.u3v_ze",
                        "Ddytkantbtitl2m.u3sf_01",
                        "smoking.u3tcigsmk",
                        "Alcohol.ethanol.u3v_za",
                        "cream",
                        "wash"))
dim(pheno)
```

# Remove remaining participants with NA's
## Estimate NA by participant

```{r}
ind.nas <- rowSums(is.na(pheno)) %>%
  data.frame(na.count = .) %>%
  bind_cols(select(pheno, NGS_ID))
```

## View
```{r}
ind.nas %>%
  arrange(desc(na.count)) %>%
  mutate(NGS_ID = factor(NGS_ID, levels = unique(NGS_ID))) %>%
  ggplot(., aes(x = NGS_ID, y = na.count)) +
  geom_bar(stat = "identity") +
  theme_linedraw()+
  ggpubr::rotate_x_text()
```

## Remove

```{r}
pheno <- pheno %>% 
  left_join(ind.nas, by = "NGS_ID") %>% 
  filter(na.count == 0)

missing_plot(pheno)
```

# Remove variables that became constant
```{r}
pheno <- remove_constant(pheno)
```

# Get Gold standards

```{r}
pheno.gold <- pheno
pheno.dic <- pheno.dic %>%
  right_join(pheno %>%
               colnames() %>% 
               data.frame(varname = .),
             by = "varname")
```
# Recode to match popgen
```{r}

pheno.gold <- pheno.gold %>% 
  mutate(smoking.u3tcigsmk = case_when(smoking.u3tcigsmk == "1" ~ "Smoker",
                                       smoking.u3tcigsmk == "2" ~ "Ex-smoker",
                                       smoking.u3tcigsmk == "3" ~ "Non-smoker"),
         gender.u3csex = str_to_sentence(gender.u3csex))
```


# Recode dictinary

```{r}

pheno.dic <- pheno.dic %>% 
  mutate(Group = case_when(varname %in% c("NGS_ID", "Participant_ID") ~ "ID",
                           varname %in% c("Site", "Agetothedeadlin.u3talter", "gender.u3csex", "BMI.u3tbmi") ~ "Anthropometrics",
                           varname %in% c("Alcohol.ethanol.u3v_za", "smoking.u3tcigsmk", "Ddytkantbtitl2m.u3sf_01") ~ "Others",
                           varname %in% c("cream", "wash") ~ "Skin care",
                           TRUE ~ Group)) %>% 
  mutate(Description = case_when(varname == "Agetothedeadlin.u3talter" ~ "Age (y)",
                                 varname == "gender.u3csex" ~ "Sex",
                                 varname == "BMI.u3tbmi" ~ "BMI (kg/m2)",
                                 varname == "Crbhydrtsabsrbb.u3v_zk" ~ "Carbohydrate (g/d)",
                                 varname == "Energy.u3v_gcal" ~ "Energy (kcal/d)",
                                 varname == "Fat.u3v_zf" ~ "Fat (g/d)",
                                 varname == "Fiber.u3v_zb" ~ "Fiber (g/d)",
                                 varname == "Protein.protein.u3v_ze" ~ "Protein (g/d)",
                                 varname == "Ddytkantbtitl2m.u3sf_01" ~ "Antibiotics (<8w)",
                                 varname == "smoking.u3tcigsmk" ~ "Smoking",
                                 varname == "Alcohol.ethanol.u3v_za" ~ "Alcohol (g/d)",
                                 varname == "cream" ~ "Lotion/cream (<24h)",
                                 varname == "wash" ~ "Wash (<24h)",
                                 varname == "Participant_ID" ~ "Participant_ID",
                                 TRUE ~ Description)) %>% 
  mutate(Variable = varname) %>% 
  mutate(Group = if_else(Group == "Anthropometrics",
                         "Age/BMI/Sex",
                         Group))
```



##Check removal
```{r}
dim(pheno.gold)
missing_plot(pheno.gold)
```
Ok, after a couple of manual tries, this seems to be the best configuration for filtering the data. This was a **manual step**.

# Check and update other dataframes
### Check resulting dimensions
```{r}
dim(pheno.gold)
```

### Get dictionary
```{r}
pheno.gold.dic <- colnames(pheno.gold) %>%
  data.frame(varname = .) %>%
  left_join(pheno.dic, by = "varname")
```

```{r}
pheno.gold.dic.values <- meta.dic.values %>%
  filter(varname %in% pheno.dic$varname)
```

## Keep microbiome data of selected samples

## Select participants
```{r}
micro.gold <- micro[match(pheno.gold$NGS_ID, rownames(micro)),]
identical(pheno.gold$NGS_ID, rownames(micro.gold))
```


## Select ASVs that are not empty
```{r}
micro.gold <- micro.gold[,colSums(micro.gold) > 0]
```

## Match tax

```{r}
tax.gold <- colnames(micro.gold) %>%
  data.frame(ASV = .) %>%
  left_join(tax, by = "ASV")
identical(colnames(micro.gold), tax.gold$ASV)
```

# Save outputs

Gold

```{r}
"pheno.all.gold.rds" %>%
  paste0(d.out, "/", .) %>%
  saveRDS(pheno.gold, .)
"micro.all.gold.rds" %>%
  paste0(d.out, "/", .) %>%
  saveRDS(micro.gold, .)
"tax.all.gold.rds" %>%
  paste0(d.out, "/", .) %>%
  saveRDS(tax.gold, .)
"pheno.dic.all.gold.rds" %>%
  paste0(d.out, "/", .) %>%
  saveRDS(pheno.gold.dic, .)
"pheno.dic.values.all.gold.rds" %>%
  paste0(d.out, "/", .) %>%
  saveRDS(pheno.gold.dic.values, .)
```


# Session information

```{r}
sessionInfo()
```

