---
title: "Filter metadata"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      code: hide
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
#bibliography: mylib.bib
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
---


# Introduction


# Preparations

## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
#v.01
# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 6,
               fig.width = 5 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)
library("rmarkdown")

# Pretty outputs
library("BiocStyle")
library("ggthemes") # add ggplot2 themes
library("ggplot2")
theme_set(theme_few(base_size = 14))

# Set seed for reproducibility
set.seed(13)
# Color blind friendly pallet
cbp1 <- c("#E69F00", "#56B4E9", "#009E73",
          "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")

scale_fill_discrete <- function(...) {
  scale_fill_manual(..., values = cbp1)
}
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = cbp1)
}

# count function
Count <- function(x){return(plyr::count(x))}
# Set output directory
d.out <- params$d.out
rm(params)
```

## Load libraries for the session

```{r}
library(ggplot2)
library(tidyverse)
library(finalfit)
library(janitor)
library(ICC)
library(vcd)
library(DataExplorer)
```

## Import data
```{r}
d.in <- "/home/lsilva/IKMB/projects/skin.associations/results/popgen/1.full.set/1.filter.samples.and.metadata"
meta <- "pheno.all.gold.rds" %>%
  paste0(d.in, "/", .) %>%
  readRDS()
micro <- "micro.all.gold.rds" %>%
  paste0(d.in, "/", .) %>%
  readRDS()
tax <- "tax.all.gold.rds" %>%
  paste0(d.in, "/", .) %>%
  readRDS()
meta.dic <- "pheno.dic.all.gold.rds" %>%
  paste0(d.in, "/", .) %>%
  readRDS()
```


# Introduction

I will filter the microbiome and the metadata to have a small, **GOLD**, selected set of samples where phenotype data is complete.

# Make new variables based on physiological parameter measured

```{r}
meta <- meta %>% 
  mutate(TEWL_specific = case_when(Site %>% as.character() == "Forehead" ~ t14544_TEWL_forehead,
                                   Site %>% as.character() == "Dorsal forearm" ~ t14546_TEWL_forearm_outside,
                                   Site %>% as.character() == "Volar forearm" ~ t14545_TEWL_forearm_inside,
                                   Site %>% as.character() == "Antecubital fossa" ~ t14547_TEWL_arm_crook),
         ph_specific = case_when(Site %>% as.character() == "Forehead" ~ t14560_pH_forehead,
                                 Site %>% as.character() == "Dorsal forearm" ~ t14562_pH_forearm_outside,
                                 Site %>% as.character() == "Volar forearm" ~ t14561_pH_forearm_inside,
                                 Site %>% as.character() == "Antecubital fossa" ~ t14563_pH_arm_crook))
```

## Add new variables to the dictionary

```{r}
to.add <- data.frame(varname = c("TEWL_specific","ph_specific"),
                     Group = "Skin",
                     Description = c(                              "Transepidermal water loss",                                                                   "pH metry")) %>% 
  mutate(
    Variable = varname,
    trait = varname,
    trait_name = varname) 

# Update dictionary
meta.dic <- meta.dic %>% 
  bind_rows(to.add) %>% 
  mutate(Description = if_else(Description == "Skin type Fitzpatrick",
                               "Skin type (Fitzpatrick)",
                               Description))
```

# Add groups to the dictionary

```{r}
# Edit Groups in dictionary

meta.dic <- meta.dic %>% 
  mutate (Group = if_else(str_detect(varname, "TEWL"),
                          "TEWL",
                          Group))
```
# Edit skin type variable due to low numbers

```{r}
meta <- meta %>% 
  mutate(t14520_Fitzpatrick_type = if_else(t14520_Fitzpatrick_type %in%
                                             c("always burns, never tans",
                                               "usually burns, tans minmally"),
                                           "always burns, never tans/usually burns, tans minmally",
                                           t14520_Fitzpatrick_type ))
```

# Keep only data of the set
Also remove redundant
```{r}

pheno <- meta %>%
  select(NGS_ID,
         Participant_ID,
         Site,
         t14110_gender,
         t14520_Fitzpatrick_type,
         t14558_TEWL_humidity,
         t14559_TEWL_temperature,
         t4_examination_age,
         t277_BMI,
         TEWL_specific,
         ph_specific)

# Sync
pheno.dic <- meta.dic %>% 
  right_join(data.frame(varname = pheno %>% colnames),
             by = "varname")
# confirming
identical(pheno %>% colnames, pheno.dic$varname)

missing_plot(pheno)
```

# Remove remaining participants with NA's
## Estimate NA by participant

```{r}
ind.nas <- rowSums(is.na(pheno)) %>%
  data.frame(na.count = .) %>%
  bind_cols(select(pheno, NGS_ID))
```

## View
```{r}
ind.nas %>%
  arrange(desc(na.count)) %>%
  mutate(NGS_ID = factor(NGS_ID, levels = unique(NGS_ID))) %>%
  ggplot(., aes(x = NGS_ID, y = na.count)) +
  geom_bar(stat = "identity") +
  theme_linedraw()+
  ggpubr::rotate_x_text()+
  geom_hline(yintercept = 10)
```

## Remove

```{r}
pheno <- pheno %>% 
  left_join(ind.nas, by = "NGS_ID") %>% 
  filter(na.count == 0) %>% 
  select(-na.count)
missing_plot(pheno)
```

# Remove variables that became constant
```{r}
pheno.gold <- remove_constant(pheno)
```

##Check removal
```{r}
dim(pheno)
missing_plot(pheno.gold)
```

# Check and update other dataframes
### Check resulting dimensions
```{r}
dim(pheno.gold)
```


### Get dictionary
```{r}
pheno.gold.dic <- colnames(pheno.gold) %>%
  data.frame(varname = .) %>%
  left_join(pheno.dic, by = "varname")
```

## Keep microbiome data of selected samples

## Select participants
```{r}
micro.gold <- micro[match(pheno.gold$NGS_ID, rownames(micro)),]
identical(pheno.gold$NGS_ID, rownames(micro.gold))
```


## Select ASVs that are not empty
```{r}
micro.gold <- micro.gold[,colSums(micro.gold) > 0]
```

## Match tax

```{r}
tax.gold <- colnames(micro.gold) %>%
  data.frame(ASV = .) %>%
  left_join(tax, by = "ASV")
identical(colnames(micro.gold), tax.gold$ASV)
```

# Save outputs

Gold

```{r}
"pheno.all.gold.rds" %>%
  paste0(d.out, "/", .) %>%
  saveRDS(pheno.gold, .)
"micro.all.gold.rds" %>%
  paste0(d.out, "/", .) %>%
  saveRDS(micro.gold, .)
"tax.all.gold.rds" %>%
  paste0(d.out, "/", .) %>%
  saveRDS(tax.gold, .)
"pheno.dic.all.gold.rds" %>%
  paste0(d.out, "/", .) %>%
  saveRDS(pheno.gold.dic, .)
```

# Session information

```{r}
sessionInfo()
```

