---
title: "Get diversity measures"
author: "Lucas Moitinho-Silva"
date: "`r Sys.Date()`"
output:
   BiocStyle::html_document:
      toc: true
      df_print: paged
      self_contained: true
      code_download: true
      highlight: tango
#bibliography: mylib.bib
editor_options:
  chunk_output_type: console
params:
  FIGPATH: "figures/"
  d.out: "./"
  index: "1"
---


# Introduction

Here, I will retrieve filter the asv table and get rarefied microbiome table, and dissimilarity matrix. Dissimilarity matrix will be calculated from rarefied data.

# Preparations

## Set global options

```{r style, echo = T, results="asis", cache=FALSE, message = F}
# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(echo = TRUE,
               tidy = FALSE,
               include = TRUE,
               fig.path = params$FIGPATH,
               dev=c("png",'pdf'),
               fig.height = 5,
               fig.width = 4 * golden_ratio,
               comment = '  ',
               dpi = 300,
               cache = FALSE)
library("rmarkdown")

# Pretty outputs
library("BiocStyle")
library("ggthemes") # add ggplot2 themes
library("ggplot2")
theme_set(theme_few(base_size = 14))

# Set seed for reproducibility
set.seed(13)

# Set output directory
d.out <- params$d.out
index <- params$index
rm(params)
```


## Load libraries for the session

```{r}
library(ggplot2)
library(tidyverse)
library(vegan)
```

```{r}
if(index == "1"){site <- "Antecubital fossa"}
if(index == "2"){site <- "Behind ear"}
```



## Import data
```{r}
d.in <- "/home/lsilva/IKMB/projects/skin.associations/results/kora/1.full.set/1.filter.samples.and.metadata"
micro <- "micro.all.gold.rds" %>%
  paste0(d.in, "/", .) %>% 
  readRDS()

pheno <- "pheno.all.gold.rds" %>%
  paste0(d.in, "/", .) %>% 
  readRDS()

pheno.dic <- "pheno.dic.all.gold.rds" %>%
  paste0(d.in, "/", .) %>% 
  readRDS()

tax <- "tax.all.gold.rds" %>%
  paste0(d.in, "/", .) %>% 
  readRDS()

```

## Sync index

```{r}


pheno <- pheno %>% 
  filter(Site == site) %>% 
  select(-Site) 



micro <- micro %>% 
  rownames_to_column("NGS_ID") %>% 
  right_join(pheno %>% select(NGS_ID), by = "NGS_ID") %>% 
  column_to_rownames("NGS_ID")

micro <- micro[,colSums(micro)>0]
tax <- tax %>% 
  right_join(micro %>%
               colnames() %>%
               data.frame( ASV = .), by = "ASV")
```


# Filter

## Set funtion to filter data

```{r}
select.asv <- function(cut.off.core, pc.cutoff, micro.st){
  
  micro.1 <- micro.st
  
  #Set cutoff core
  min.participants <- nrow(micro.1) * cut.off.core
  print(min.participants)
  # Get ASVs that pass the core cutoff
  core <- micro.1 > 0
  core <- core %>% colSums()
  core <- core[core > min.participants]
  core <- names(core)
  # Set cutoff abundance
  abun <- colSums(micro.1)
  total.seqs <- sum(abun)
  min.abun <- (pc.cutoff * total.seqs) / 100
  print(min.abun)
  # Get ASVs that pass abundance cutoff
  abun <- colSums(micro.1)
  abun <- abun[abun > min.abun]
  abun <- names(abun)
  # Combine the results
  asv.to.keep.1 <- intersect(core, abun) %>% unique()
  
  d <- data.frame(ASV = asv.to.keep.1, stringsAsFactors = F) 
  return(d)}
```

## Set cut-off and filter
```{r}

cut.off.core <- 0 #proportion
# Abundance
pc.cutoff <- 0.1 #percentage

# rarefy for filter
min.count <- 5000
micro.r <- rrarefy(micro, min.count) %>% 
  data.frame()
micro.r <-  micro.r[, colSums(micro.r) > 0]
asvs <- select.asv(cut.off.core, pc.cutoff, micro.r) %>% 
  mutate(status = "Bona fidi")

micro.filtered <- micro[, colnames(micro) %in% asvs$ASV]
micro.filtered <- micro.filtered[rowSums(micro.filtered) > 5000, ]

# Gather stats

stats <- data.frame(Stage = "Before",
                    seqs = micro %>% colSums() %>% sum(),
                    samples = nrow(micro))
stats <- bind_rows(stats, 
                   data.frame(Stage = "After",
                    seqs = micro.filtered %>% colSums() %>% sum(),
                    samples = nrow(micro.filtered)))

paste0("i", index, ".", "stats.filtering.tsv") %>%
  paste0(d.out, "/", .) %>% 
  write_tsv(stats, .)
# Rename
micro <- micro.filtered

# remove temporary files
rm(micro.r, micro.filtered)
```

# Sync Files

```{r}
tax <- tax %>%
  right_join(data.frame(ASV = colnames(micro)), by  ="ASV")

pheno <- pheno %>%
  right_join(data.frame(NGS_ID = rownames(micro)), by  = "NGS_ID")
```


# Rarefy data

```{r}
min.count <- 5000
micro.r <- rrarefy(micro, min.count) %>% 
  data.frame()
micro.r <-  micro.r[, colSums(micro.r) > 0]
```

# Get distances

```{r}
micro.dis <- micro.r %>%
  vegdist("bray")
```


# Saving output

```{r}
paste0("i", index, ".", "micro.rarefied.rds") %>% 
  paste0(d.out, "/", .) %>% 
  saveRDS(micro.r, .)

paste0("i", index, ".","micro.dis.rds") %>% 
  paste0(d.out, "/", .) %>% 
  saveRDS(micro.dis, .)

paste0("i", index, ".","pheno.rds") %>% 
  paste0(d.out, "/", .) %>% 
  saveRDS(pheno, .)

paste0("i", index, ".", "micro.rds") %>% 
  paste0(d.out, "/", .) %>% 
  saveRDS(micro, .)

paste0("i", index, ".", "pheno.dic.rds") %>% 
  paste0(d.out, "/", .) %>% 
  saveRDS(pheno.dic, .)

paste0("i", index, ".", "tax.rds") %>% 
  paste0(d.out, "/", .) %>% 
  saveRDS(tax, .)

```


# Session information

```{r}
sessionInfo()
```
